Guided Proofreading of Automatic Segmentations for Connectomics

### 2016-11-13 paper/introduction.tex
Intro.

In connectomics, neuroscientists annotate neurons and their connectivity within 3D volumes to gain insight into the functional structure of the brain. Rapid progress in automatic sample preparation and electron microscopy (EM) acquisition techniques has made it possible to image large volumes of brain tissue at $\approx4\, nm$ per pixel to identify cells, synapses, and vesicles. For $40\, nm$ thick sections, a $1\, mm^3$ volume of brain contains $10^{15}$ voxels, or 1 petabyte of data. With so much data, manual annotation is infeasible, and automatic annotation methods are needed ~\cite{jain2010,Liu2014,GALA2014,kaynig2015large}.

@@

Our goal is to add automatic detection of split and merge errors to proofreading tools. We design automatic classifiers that detect split and merge errors in 2D segmentations so the user does not need to visually inspect the whole data volume to spot errors. A proofreading tool then recommends regions with a high probability of an error to the user, and suggest corrections to accept or reject. We call this process \textit{guided proofreading} (GP).

As our main contribution, we introduce classifiers to detect merge- and split errors based on a convolutional neural network (CNN). We believe that this is the first time that deep learning is applied to the task of proofreading. Our classifiers work on top of any existing automatic segmentation method to find potential errors and suggest corrections. Given a membrane segmentation from a fast automatic method, our classifiers operate on the boundaries of whole cell regions. Compared to techniques that must analyze every input pixel, we reduce the data analysis to the boundaries only. First, we train a CNN to detect only split errors. The output of this network is a probability whether a boundary between two segments is valid or not. We then reuse the same network to also detect merge errors by generating possible boundaries within a cell and inverting the split error score. We create corrections for both types of errors which can be accepted or rejected. This reduces the proofreading operation to simple yes/no decisions.

We further propose a greedy algorithm to perform proofreading automatically. Possible erroneous regions are sorted by their score and the algorithm sequentially corrects merge errors and split errors until a configurable threshold is reached. If ground truth data is available, we can use a selection oracle to drive this algorithm. The oracle only accepts corrections which improve the automatic segmentation.

This way,

 and measure performance as the adapted Rand error which is a common metric for segmentation comparison~\cite{RAND}. 
Our system is integrated into an existing proofreading workflow for large connectomics data. For this, we also explore an active label suggestion approach in addition to the ranking obtained by guided proofreading. We quantitatively validate automatic and human-driven variations of guided proofreading on five different real-world connectomics datasets of mouse as well as fruitfly (drosophila) brain. To study the performance of novice and expert proofreaders, we perform a between-subjects experiment and ask participants to proofread a publicly available dataset. For comparison, we establish two baselines: a recently published fully interactive proofreading tool named \textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014} and semi-automatic \textit{focused proofreading} (FP) approach by Plaza~\cite{focused_proofreading}. In all experiments, we significantly outperform both interactive proofreading as well as Plaza's method.

As a consequence, we are able to provide tools to proofread segmentations more efficiently, and so better tackle large volumes of connectomics imagery.

>>>>

In connectomics, neuroscientists annotate neurons and their connectivity within 3D volumes to gain insight into the functional structure of the brain. Rapid progress in automatic sample preparation and electron microscopy (EM) acquisition techniques has made it possible to image large volumes of brain tissue at $\approx4\, nm$ per pixel to identify cells, synapses, and vesicles. For $40\, nm$ thick sections, a $1\, mm^3$ volume of brain contains $10^{15}$ voxels, or 1 petabyte of data. With so much data, manual annotation is infeasible, and automatic annotation methods are needed~\cite{jain2010,Liu2014,GALA2014,kaynig2015large}.

@@

Our goal is to add automatic detection of split and merge errors to proofreading tools. We design automatic classifiers that detect split and merge errors in 2D segmentations so the user does not need to visually inspect the whole data volume to spot errors. A proofreading tool then recommends regions with a high probability of an error to the user, and suggest corrections to accept or reject. We call this process \textit{guided proofreading}.

In this paper, we introduce classifiers to detect merge- and split errors based on a convolutional neural network (CNN). We believe that this is the first time that deep learning is applied to the task of proofreading. Our classifiers work on top of any existing automatic segmentation method to find potential errors and suggest corrections. Given a membrane segmentation from a fast automatic method, our classifiers operate on the boundaries of whole cell regions. Compared to techniques that must analyze every input pixel, we reduce the data analysis to the boundaries only. First, we train a CNN to detect only split errors. The output of this network is a probability whether a boundary between two segments is valid or not. We then reuse the same network to also detect merge errors by generating possible boundaries within a cell and inverting the split error score. We create corrections for both types of errors which can be accepted or rejected. This reduces the proofreading operation to simple yes/no decisions.

We further propose a greedy algorithm to perform proofreading. Possible erroneous regions are sorted by their score and the algorithm iteratively suggests a correction for each region. A user then works through this stream of regions and corrections. In a  forced choice setting, the user chooses one of two possibilities to advance. This choice can be also performed automatically by running the algorithm until a configurable threshold is reached. In addition, if ground truth data is available, we can use a selection oracle to drive the forced choice selection. The oracle only accepts corrections which improve the automatic segmentation. This equals perfect proofreading.

We evaluate our method automatically using a threshold and a selection oracle on multiple real-world connectomics datasets. To evaluate the forced choice setting, we perform a quantitative user study. The study targets non-experts with no previous experience of proofreading electron microscopy data. We ask the participants to proofread a small segmentation volume in a fixed time frame by performing yes/no decisions. The user study is designed as a between-subjects experiment and compares guided proofreading against two other methods: a recently published fully interactive proofreading tool named \textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014} and semi-automatic \textit{focused proofreading} approach by Plaza~\cite{focused_proofreading}. We also asked four domain experts to use guided proofreading and focused proofreading for additional comparison.

Our first contribution is a classifier for split error detection based on a convolutional neural network. The classifier performs well even when trained with little amounts of training data. This is important since generating ground truth labels in connectomics is done by manually labeling pixels which is very time-consuming. Our second contribution is a mechanism to identify merge-errors by re-using our split error classifier. Merge errors are less common than split errors in the usually oversegmented automatic labelings. However, interactive correction of merge errors by drawing split lines is more cumbersome than correcting split errors. Our method reduces this to a single click by providing the potential correction. The split and merge error identification is executed as a greedy algorithm to perform corrections of segmentation volume, the third contribution of this paper. The algorithm can be driven automatically with a threshold, by an oracle based on ground truth and interactively in a forced choice setting. Our final contribution is our quantitative user study. We present statistically significant results showing that novice and expert users of guided proofreading are able to proofread a given dataset better and faster than with existing interactive and semi-automatic proofreading tools. As a consequence, we are able to provide tools to proofread segmentations more efficiently, and so better tackle large volumes of connectomics imagery.
